---
title: "Organize voucher directories"
format: html
execute:
  echo: false
  eval: true
  freeze: auto
  cache: false
  warning: false
  message: false
  include: true
  inline: true
---

```{r setup, include=FALSE}
library(forplotR)
```

This guide demonstrates how to use the `mk_voucher_dirs` function in the `forplotR` package to create and organize image directories of vouchers recorded in plot censuses based on their **updated taxonomic identification**.

## Function Overview

The `mk_voucher_dirs()` function organizes **voucher image folders** according to the most up-to-date taxonomic identification found in a **ForestPlots-style** Excel file.

It is safe to run multiple times:

* **First run:** creates the main output folder (default: `"voucher_imgs"`) and, for each voucher marked as **Collected**, creates the directory structure `Family/Genus/Voucher`.
* **Subsequent runs (reorganization):** detects vouchers placed under outdated genera and **moves images** to the correct folder **without overwriting** existing files; then **removes empty, outdated directories**.

## Arguments

| Argument       | Description                                                                                      |
| -------------- | ------------------------------------------------------------------------------------------------ |
| `fp_file_path` | Path to the Excel file (ForestPlots template) containing voucher data. **Required**.             |
| `output_dir`   | Directory where the voucher folder structure will be created/updated. Default: `"voucher_imgs"`. |

## Basic Use Case

```r
mk_voucher_dirs(
  fp_file_path = "data/RUS_plot.xlsx",
  output_dir   = "voucher_imgs"
)
```

## Reorganizing After Taxonomic Updates

If some vouchers were reidentified (e.g., genus changed), just run the function again:

```r
mk_voucher_dirs(
  fp_file_path = "data/RUS_plot.xlsx",
  output_dir   = "voucher_imgs"
)
```

* Images found under an outdated genus folder are **moved** to the correct genus folder.
* The function **does not overwrite** existing files in the correct destination.
* Any **empty old voucher folders** are removed.

## Using a Custom Output Directory

```r
mk_voucher_dirs(
  fp_file_path = "data/Inventory_A.xlsx",
  output_dir   = "parcelA_voucher_imgs"
)
```

## Inspecting the Resulting Structure

```r
# List created directories (first levels)
list.dirs("voucher_imgs", recursive = FALSE)

# Inspect all voucher-level directories
dirs <- list.dirs("voucher_imgs", recursive = TRUE, full.names = TRUE)
head(dirs, 20)
```

```{r}
#| echo: false
# Quick check: count voucher directories under output_dir
out <- "voucher_imgs"
if (dir.exists(out)) {
  all_dirs <- list.dirs(out, recursive = TRUE, full.names = TRUE)
  voucher_dirs <- all_dirs[grepl("[A-Za-z]+/[A-Za-z]+/[A-Za-z]{2,}[0-9]{3,}", all_dirs)]
  cat("Total voucher directories found:", length(voucher_dirs), "
")
} else {
  cat("Run mk_voucher_dirs() first to create", out, "
")
}
```

## Tips

* Ensure your Excel file includes columns like `Family`, `Original determination`, `Voucher`, and `Collected`.
* The **genus** is derived from the **first token** in `Original determination` (text before the first space).
* Re-run the function whenever identifications change; it is designed to be **idempotent**.
* Consider keeping a **backup** before large reorganizations when working with many images.
* If a destination folder already has images, the function **skips moving** to avoid overwriting—review manually if you want to merge.

## See Also

* `plot_for_balance()`: Parcel report with maps, checklist, navigation, and optional AGB.
* `plot_html_map()`: Interactive HTML maps for plots.
* `fp_herb_converter()`: Convert ForestPlots sheets to herbarium formats.
